---
title: "Video Game Trends: 2000-2010"
author: "Ryan"
execute:
    echo: FALSE
    message: FALSE
    warning: FALSE
---

<h2 style="text-align: center;">Project Summary</h2>
<br>
Video game popularity and sales changed dramatically between 2000 and 2010. I gathered data from a Kaggle dataset containing global sales information on thousands of video games released during that decade. After cleaning and filtering the data, I focused on top-selling titles and major platforms across different regions. 






```{r setup}
library(tidyverse)
library(plotly)
library(ggplot2)

vgsales <- read_csv("vgsales.csv")

top10_games <- vgsales %>%
  filter(Year >= 2000, Year <= 2010) %>%  # keep only games from 2000-2010
  arrange(desc(Global_Sales)) %>%          # sort by Global_Sales descending
  slice_head(n = 10)                       # pick top 10 rows


```





```{r}
p <- ggplot(top10_games, aes(
  x = Name,
  y = Global_Sales,
  fill = Platform,
  text = paste0("Year: ", Year, "<br>",
                "Global Sales: ", Global_Sales, " million<br>",
                "Platform: ", Platform)
)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 10 Best-Selling Video Games (2000-2010)",
    x = "Video Game",
    y = "Global Sales (Millions)",
    fill = "Platform"
  ) +
  theme_minimal()

ggplotly(p, tooltip = "text")

```
<br><br>

The highest-selling games during this time were predominantly released on Nintendo’s DS and Wii, showcasing their strong market dominance. Genres like action and sports consistently led global sales, shaping the gaming industry’s growth. 
```{r}
library(tidyverse)
library(plotly)

# Assuming vgsales and top10_games are already loaded:
top10_games <- vgsales %>%
  filter(Year >= 2000, Year <= 2010) %>%
  arrange(desc(Global_Sales)) %>%
  slice_head(n = 10)

# Summarize sales by region
continent_sales <- top10_games %>%
  summarise(
    North_America = sum(NA_Sales, na.rm = TRUE),
    Europe = sum(EU_Sales, na.rm = TRUE),
    Japan = sum(JP_Sales, na.rm = TRUE),
    Other = sum(Other_Sales, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = everything(), names_to = "Region", values_to = "Sales")

# Interactive pie chart with centered title and top margin
plot_ly(
  data = continent_sales,
  labels = ~Region,
  values = ~Sales,
  type = "pie",
  textinfo = "label+percent",
  hoverinfo = "text",
  text = ~paste("Region:", Region, "<br>Sales:", round(Sales, 2), "million"),
  marker = list(colors = c("red", "green", "blue", "orange"))
) %>%
  layout(
    title = list(
      text = "Top 10 Games Sales by Region (2000–2010)",
      x = 0.43,
      xanchor = "center",
      font = list(size = 18)
    ),
    margin = list(t = 80)  # Adjust top margin for spacing
  )

```
To better understand the worldwide impact, this project includes an interactive pie chart showing the regional distribution of video game sales, offering insight into how different markets contributed to overall sales throughout the decade.
