---
title: "Video Game Trends: 2000-2010"
author: "Ryan"
execute:
    echo: FALSE
    message: FALSE
    warning: FALSE
---

<h2 style="text-align: center;">Project Summary</h2>
<br>
Video game popularity and sales changed dramatically between 2000 and 2010. I gathered data from a Kaggle dataset containing global sales information on thousands of video games released during that decade. After cleaning and filtering the data, I focused on top-selling titles and major platforms across different regions. Through visualizations like bar charts and pie charts, I was able to explore which consoles dominated, how regional preferences differed, and which genres led global trends. Nintendo’s DS and Wii platforms stood out as major contributors to overall sales, particularly in North America and Japan. This analysis helped uncover not only which games were most successful, but also how market dynamics and consumer behavior varied around the world.


<hr style="border: none; border-top: 1px solid #ccc; margin: 1em 0;">




```{r setup}
library(tidyverse)
library(plotly)
library(ggplot2)
library(webshot2)

vgsales <- read_csv("vgsales.csv")

top10_games <- vgsales %>%
  filter(Year >= 2000, Year <= 2010) %>%  # keep only games from 2000-2010
  arrange(desc(Global_Sales)) %>%          # sort by Global_Sales descending
  slice_head(n = 10)                       # pick top 10 rows


```





```{r}
library(tidyverse)
library(plotly)

# Load and filter top 10 games
top10_games <- vgsales %>%
  filter(Year >= 2000, Year <= 2010) %>%
  arrange(desc(Global_Sales)) %>%
  slice_head(n = 10)

# Create the bar chart with Plotly
p <- ggplot(top10_games, aes(
  x = reorder(Name, Global_Sales),
  y = Global_Sales,
  fill = Platform,
  text = paste0("Year: ", Year, "<br>",
                "Global Sales: ", Global_Sales, " million<br>",
                "Platform: ", Platform)
)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 10 Best-Selling Video Games (2000–2010)",
    x = "Video Game",
    y = "Global Sales (Millions)",
    fill = "Platform"
  ) +
  theme_minimal()

# Convert to interactive Plotly chart with transparent background
ggplotly(p, tooltip = "text") %>%
  layout(
    title = list(
      text = "Top 10 Best-Selling Video Games (2000–2010)",
      x = 0.37,
      xanchor = "center"
    ),
    paper_bgcolor = 'rgba(0,0,0,0)',  # full background transparent
    plot_bgcolor = 'rgba(0,0,0,0)'    # plot panel background transparent
  )

```
<br><br>

The highest-selling games during this time were predominantly released on Nintendo’s DS and Wii, showcasing their strong market dominance. Genres like action and sports consistently led global sales, shaping the gaming industry’s growth. 
```{r}
# Summarize sales by region
continent_sales <- top10_games %>%
  summarise(
    North_America = sum(NA_Sales, na.rm = TRUE),
    Europe = sum(EU_Sales, na.rm = TRUE),
    Japan = sum(JP_Sales, na.rm = TRUE),
    Other = sum(Other_Sales, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = everything(), names_to = "Region", values_to = "Sales")

# Create interactive pie chart
plot_ly(
  data = continent_sales,
  labels = ~Region,
  values = ~Sales,
  type = "pie",
  textinfo = "label+percent",
  hoverinfo = "text",
  text = ~paste("Region:", Region, "<br>Sales:", round(Sales, 2), "million"),
  marker = list(colors = c("red", "green", "blue", "orange"))
) %>%
  layout(
    title = list(
      text = "Top 10 Game Sales by Region (2000–2010)",
      x = 0.44,
      xanchor = "center",
      font = list(size = 18)
    ),
    margin = list(t = 80),
    paper_bgcolor = 'rgba(0,0,0,0)',  # Transparent background
    plot_bgcolor = 'rgba(0,0,0,0)'    # Transparent plot panel
  )
```
To highlight the global impact, this project features an interactive pie chart showing the regional distribution of the top 10 video game sales, revealing how various markets contributed to total sales throughout the decade.

